---
title: "Collusion Project"
output: html_notebook
---

---
title: "Ohio School Milk Collusion"
output: html_notebook
---

```{r}
rm(list = ls())
library(tidyverse)
library(lubridate)
library(party)
```

```{r}
District <- readr::read_table("DISTRICT.PRN", col_names = FALSE, na = "NA") # file does not have headers
MilkData <- read_csv("ohio_milk_data.csv")
```

## Clean data from District.PRN
```{r}
# Find individual district IDs in rows X4-X7 and combine them into a single district id
District <-
  District %>%
  mutate(D4 = str_extract(string = X4, pattern = "^0{1}\\d{5}"),
         D5 = str_extract(string = X5, pattern = "^0{1}\\d{5}"),
         D6 = str_extract(string = X6, pattern = "^0{1}\\d{5}"),
         D7 = str_extract(string = X7, pattern = "^0{1}\\d{5}"),
         district_id = coalesce(D4, D5, D6, D7)) %>%
  select(!c(D4,D5,D6,D7))

# Create a single variable for city
District <-
  District %>%
  unite(city, c(X2, X3), sep = " ", remove = TRUE)



```

## Add city name to ohio_milk data
```{r}
# Isolate district ID in MilkData
MilkData <- 
  MilkData %>%
  separate(contract_id, c("year", "district_id"), remove = FALSE) # keeping original to bid ID purposes

# Join District and MilkData so that city names are associated with district IDs
MilkData <- 
  MilkData %>%
  left_join(District %>% select(district_id, city), by = "district_id")

# Clean up MilkData
MilkData <- 
  MilkData %>%
  select(!c(...1, inflation_adjusted_bid, inflation_raw_milk_price_adjusted_bid)) %>%
  select(contract_id, year, city, district_id, bid, firm, winning_bid, is_cartel_firm, only_cartels_bidding_on_contract, number_of_bidders_on_contract) %>%
  mutate(district_id = as.numeric(district_id),
         year = as.numeric(year))
```

## Summary Statistics
```{r}
summary(MilkData)
```

```{r}
## Number of districts
MilkData %>%
  count(district_id)

# Number of contracts where there were at least 5 bidders.
many_bidders <-
  MilkData %>%
  group_by(contract_id) %>%
  summarize(total_bids = n()) %>%
  arrange(desc(total_bids)) %>%
  filter(total_bids >= 5)

#Find number of cartel members per contract
cartel_members_per_contract <-
  MilkData %>%
  filter(is_cartel_firm == 1) %>%
  group_by(contract_id) %>%
  summarize(cartel_per_contract = n())

## Add number of cartel members bidding on a contract to MilkData
MilkData <-
  MilkData %>%
  left_join(cartel_members_per_contract, by = c("contract_id" = "contract_id"))

# Add cartel_members_per_contract information to many_bidders
many_bidders <-
  many_bidders %>%
  left_join(MilkData, by = c("contract_id" = "contract_id")) %>%
  mutate(year = as.numeric(year)) %>%
  coalesce()

```
* There are only 25 occasions where there were at least 25 bids on a single contract for a given year.

## Summary statistics for contracts with >5 bidders
```{r}
summary(many_bidders)

# Collusive vs non-collusive bidders
many_bidders %>%
  group_by(is_cartel_firm) %>%
  summarize(total_bids = n())

# Contracts with collusive bidder vs non-collusive contracts
many_bidders %>%
  group_by(contract_id) %>%
  summarize(cartel_percent = cartel_per_contract/number_of_bidders_on_contract) %>%
  distinct(contract_id, .keep_all = TRUE) %>%
  ungroup() %>%
  mutate(cartel_percent = ifelse(is.na(cartel_percent), 0, cartel_percent)) %>%
  arrange(desc(cartel_percent))
```

## Coefficent of Variation 
The coefficient of variation, $\large CV_j$ is the ratio of the standard deviation of the bid to the mean for each contract $\large j$. The lower the value of $\large CV_j$, the more suspicious the behavior.

```{r}
CV_j <- 
  many_bidders %>%
  select(contract_id, bid) %>%
  group_by(contract_id) %>%
  mutate(bid_name = "bid",
         brank = rank(bid, ties.method = "random")) %>%
  unite(bid_rank, c(bid_name, brank), sep = "_", remove = TRUE ) %>%
  pivot_wider(names_from = bid_rank, values_from = bid) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(sd = sd(c(bid_1,bid_2,bid_3, bid_4, bid_5, bid_6), na.rm = TRUE),
         mu = mean(c(bid_1,bid_2,bid_3, bid_4, bid_5, bid_6), na.rm = TRUE),
         CV = sd/mu) %>%
  left_join(many_bidders %>% select(contract_id, cartel_per_contract), by = "contract_id")

```

## Relative Distance




